---
title: "Midpoint Deliverable"
author: "Kai Jeffreys"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)

df <- read.csv("data/NHIS-Data.csv")

source("scripts/summary_info_list.R")
source("scripts/summary_info_table.R")
source("scripts/race_and_CV_chart.R")
source("scripts/Risk_factors_comparison.R")
source("scripts/cv_and_risks_byAge_byRegion.R")
```

## Introduction
For this assignment, we are using the NHIS Data set. The data was collected 
through personal household interviews, where participants are asked a series of 
questions about their personal health, including their heart. Data was collected
by the NHIS, a branch of the CDC. The data shows the proportion of the people
who surveyed had cardiovascular disease, among other risk factors, within in a 
certain subset of the overall population of people surveyed. The proportion is
shown in the Data_Value column as a percent. The data is sorted into breakout
groups, like age, gender, or race. Since the breakout groups are all in one 
column, it is hard to make an easy summary of the data, since there is so much 
overlapping. It also has different columns for year and location, so there are 
more ways to break down the data other than the breakout groups.

## Summary Information
```{r, echo=FALSE, message=FALSE}
summary_info_list <- create_summary_info_list(df)
```
Since **`r summary_info_list$avg_risk_percent_us`**% of American's 
are at risk of cardiovascular diseases and **`r summary_info_list$avg_diseases_percent_us`**
% have cardiovascular diseases, heart issues are a top health concern. **`r summary_info_list$most_prevalent_risk`**
is the most prevalent risk and **`r summary_info_list$most_prevalent_disease`** 
the most prevalent cardiovascular disease. Between 2000 and 2012 both risk and cardiovascular diseases peeked in 
**`r summary_info_list$highest_diseases_year`**. The housing market 
crashing then increased stress levels and would have influenced risks and 
diseases peeking.

## Summary Table
```{r, echo = FALSE, message=FALSE}
kable(create_summary_table(df))
```

The breakout column had three dimensions: age, gender, and race. We picked 
gender, so we didn't have overlays with the data values. Males tend to have a 
higher average percent of cardiovascular diseases than females except for the 
most prevalent one, hypertension. Females also report a higher percent of 
inactivity, but a lower percent of smoking. Males have almost double acute 
myocardial infarction (heart attack) than females. The other cardiovascular 
disease that males and females have significant differences in is coronary heart 
disease. This makes sense since a complete blockage from coronary heart disease
can lead to a heart attack. 

## Cardiovascular Disease and Race
```{r, echo=FALSE, message=FALSE}
bar_chart(df)
```

This bar chart helps show a relationship between race and the chance of having 
major cardiovascular disease in the entirety of the US. Along with showing the 
mean, the chart also shows the average low and high estimates. This is because showing just the mean might not be entirely reliable, but having the low and 
high estimates show the range where the true average percent of each race should
lie. 

Information chart reveals:

- That there is a difference in CV rates among different races. 
- Non-Hispanic Asians have the lowest rates, followed by Non-Hispanic Whites. 
- Non-Hispanic Blacks and others have clearly the highest rates, with others 
having the highest by a significant margin (Other races include Native Americans
, Pacific Islander, etc).

## Risk Factors Comparison
```{r, echo=FALSE, message=FALSE}
create_line_graph(df)
```

This chart shows the relationships between the year and the percent of people
who have or have used particular risk factors for cardiovascular disease. The 
chart also breaks the data down for separate age groups.

The chart shows:

- The percent of people decreases for both being physically inactive and smoking
as time goes on as we see the percent decreasing as the years get larger
- The older a person gets, the more likely they are to be physically inactive,
since the lines that include more elder people are the ones that are shown to 
have a higher percent
- The younger a person gets, the more likely the are to smoke, since the colors
that include the young adults have higher percentages of smoking

## Cardiovascular Diseases by Age and Region
```{r, echo=FALSE, message=FALSE, warning=FALSE}
cv_risks_age_and_region(df)
```

This attempts to understand the relationships between age, CV risk factors and
the prevalence of CV Disease. It also attempts to see if these variables are 
related differently in various geographical regions of the US from the years 
2000 to 2017.

Insights from the data include:

- Regardless of age or region, people, medially, acquire cardiovascular (CV)
diseases 
at a smaller rate than people exhibit the risk factors (low physical activity 
and smoking) in the US.
- Older people correspond to a higher prevalence of CV risk factors and CV
diseases in the US. 
- The prevalence of CV risk factors before 64 years has q distribution with 
outlying data points, but after 64 years, the data is more spread out with more 
higher prevalence rates in the 4rd inter-quartile range of the box plots. This 
is consistent with the jump in prevalence of CV diseases after 64 years of age.
This strong correlation may be due to 64 being the retirement age in US. 
- There are no significant differences in CV risk factors or diseases between
various regions of the US.
